---
title: Analysing Crimes and Outcomes in the West Midlands between January and August 2020
author: "Vanessa Fillis"
date: "10/15/2020"
output: html_document
---

# Analysing the crimes and outcomes of reported crimes in the West Midlands between January 2020 and August 2020

At first, download the data from https://data.police.uk/data/ and save the .csv-files in your working directory. 

## Import all reported crimes

Then, import all reported crimes using **read.csv()**

```{r}
crimesWestMidlands0120 <- read.csv("2020-01-west-midlands-street.csv", stringsAsFactors = FALSE)
crimesWestMidlands0220 <- read.csv("2020-02-west-midlands-street.csv", stringsAsFactors = FALSE)
crimesWestMidlands0320 <- read.csv("2020-03-west-midlands-street.csv", stringsAsFactors = FALSE)
crimesWestMidlands0420 <- read.csv("2020-04-west-midlands-street.csv", stringsAsFactors = FALSE)
crimesWestMidlands0520 <- read.csv("2020-05-west-midlands-street.csv", stringsAsFactors = FALSE)
crimesWestMidlands0620 <- read.csv("2020-06-west-midlands-street.csv", stringsAsFactors = FALSE)
crimesWestMidlands0720 <- read.csv("2020-07-west-midlands-street.csv", stringsAsFactors = FALSE)
crimesWestMidlands0820 <- read.csv("2020-08-west-midlands-street.csv", stringsAsFactors = FALSE)
```

## Join the data frames 

Join all individual data frames using **rbind()** and store the result in a new variable. 

```{r}
crimesWestMidlands0120to0820 <- rbind(crimesWestMidlands0120, crimesWestMidlands0220, crimesWestMidlands0320, crimesWestMidlands0420, crimesWestMidlands0520, crimesWestMidlands0620, crimesWestMidlands0720, crimesWestMidlands0820)
```

## Explore the new data frame

How many crimes have been reported?

```{r}
length(crimesWestMidlands0120to0820$Crime.type)
```

Which crime type has been reported the most?

```{r}
crimetypesWestMidlands2020 <- table(crimesWestMidlands0120to0820$Crime.type)
crimetypesWestMidlands2020 <- crimetypesWestMidlands2020[order(-crimetypesWestMidlands2020)]
crimetypesWestMidlands2020
```


## Import the outcomes 

Next, import all outcomes: 

```{r}
outcomesWestMidlands0120 <- read.csv("2020-01-west-midlands-outcomes.csv", stringsAsFactors = FALSE)
outcomesWestMidlands0220 <- read.csv("2020-02-west-midlands-outcomes.csv", stringsAsFactors = FALSE)
outcomesWestMidlands0320 <- read.csv("2020-03-west-midlands-outcomes.csv", stringsAsFactors = FALSE)
outcomesWestMidlands0420 <- read.csv("2020-04-west-midlands-outcomes.csv", stringsAsFactors = FALSE)
outcomesWestMidlands0520 <- read.csv("2020-05-west-midlands-outcomes.csv", stringsAsFactors = FALSE)
outcomesWestMidlands0620 <- read.csv("2020-06-west-midlands-outcomes.csv", stringsAsFactors = FALSE)
outcomesWestMidlands0720 <- read.csv("2020-07-west-midlands-outcomes.csv", stringsAsFactors = FALSE)
outcomesWestMidlands0820 <- read.csv("2020-08-west-midlands-outcomes.csv", stringsAsFactors = FALSE)
```

## Join all outcomes using rbind()

Join all individual data frames using **rbind()** and store the result in a new variable. 

```{r}
outcomesWestMidlands0120to0820 <- rbind(outcomesWestMidlands0120, outcomesWestMidlands0220, outcomesWestMidlands0320, outcomesWestMidlands0420, outcomesWestMidlands0520, outcomesWestMidlands0620, outcomesWestMidlands0720, outcomesWestMidlands0820)
```

## Explore the data frame

How many outcomes have been published? 

```{r}
length(outcomesWestMidlands0120to0820$Outcome.type)
```

How many times does each outcome type occur?

```{r}
outcometypesWestMidlands2020 <- table(outcomesWestMidlands0120to0820$Outcome.type)
outcometypesWestMidlands2020 <- outcometypesWestMidlands2020[order(-outcometypesWestMidlands2020)]
outcometypesWestMidlands2020
```

## Join crimes and outcomes 

Join all crimes and outcomes by their crime.id using **merge()**.

When merging the data frames, only the rows where the crime IDs in the data frames *crimesWestMidlands0120to0820* and *outcomesWestMidlands0120to0820* match, are stored in the new data frame. 

This means that crimes without a crime ID or crimes that have not been solved yet, are not stored in the new data set. 

```{r}
crimesoutcomesWestMidlands2020 <- merge(crimesWestMidlands0120to0820, outcomesWestMidlands0120to0820, by = "Crime.ID")
```

## Explore the new data frame

```{r}
summary(crimesoutcomesWestMidlands2020)
```

## Crime outcomes for crimes reported in the West Midlands between January and August 2020

Find out what the most common crime outcome for the crimes reported in the West Midlands in 2020 was by using the **table()** function.

```{r}
crimesoutcomesWM2020table <- table(crimesoutcomesWestMidlands2020$Outcome.type)
crimesoutcomesWM2020table <- crimesoutcomesWM2020table[order(-crimesoutcomesWM2020table)]
crimesoutcomesWM2020table
#save as .csv
write.csv(crimesoutcomesWM2020table, file = "crimesandoutcomesWM2020table.csv")
```

## Proportions of total number of reported crimes in that period

What proportion does each outcome type have of the overall number of reported crimes in the West Midlands between January 2020 and August 2020? 

```{r}
crimesoutcomesWM2020table[1]/length(crimesWestMidlands0120to0820$Crime.type)
crimesoutcomesWM2020table[2]/length(crimesWestMidlands0120to0820$Crime.type)
crimesoutcomesWM2020table[3]/length(crimesWestMidlands0120to0820$Crime.type)
crimesoutcomesWM2020table[4]/length(crimesWestMidlands0120to0820$Crime.type)
crimesoutcomesWM2020table[5]/length(crimesWestMidlands0120to0820$Crime.type)
crimesoutcomesWM2020table[6]/length(crimesWestMidlands0120to0820$Crime.type)
crimesoutcomesWM2020table[7]/length(crimesWestMidlands0120to0820$Crime.type)
crimesoutcomesWM2020table[8]/length(crimesWestMidlands0120to0820$Crime.type)
crimesoutcomesWM2020table[9]/length(crimesWestMidlands0120to0820$Crime.type)
crimesoutcomesWM2020table[10]/length(crimesWestMidlands0120to0820$Crime.type)
crimesoutcomesWM2020table[11]/length(crimesWestMidlands0120to0820$Crime.type)
```

## Sum of all outcome proportions  

For what proportion of the reported crimes is an outcome available?

```{r}
sum(crimesoutcomesWM2020table[1]/length(crimesWestMidlands0120to0820$Crime.type),
crimesoutcomesWM2020table[2]/length(crimesWestMidlands0120to0820$Crime.type),
crimesoutcomesWM2020table[3]/length(crimesWestMidlands0120to0820$Crime.type),
crimesoutcomesWM2020table[4]/length(crimesWestMidlands0120to0820$Crime.type),
crimesoutcomesWM2020table[5]/length(crimesWestMidlands0120to0820$Crime.type),
crimesoutcomesWM2020table[6]/length(crimesWestMidlands0120to0820$Crime.type),
crimesoutcomesWM2020table[7]/length(crimesWestMidlands0120to0820$Crime.type),
crimesoutcomesWM2020table[8]/length(crimesWestMidlands0120to0820$Crime.type),
crimesoutcomesWM2020table[9]/length(crimesWestMidlands0120to0820$Crime.type),
crimesoutcomesWM2020table[10]/length(crimesWestMidlands0120to0820$Crime.type),
crimesoutcomesWM2020table[11]/length(crimesWestMidlands0120to0820$Crime.type))
```

For the rest no outcome is available: 
 
```{r}
1-sum(crimesoutcomesWM2020table[1]/length(crimesWestMidlands0120to0820$Crime.type),
crimesoutcomesWM2020table[2]/length(crimesWestMidlands0120to0820$Crime.type),
crimesoutcomesWM2020table[3]/length(crimesWestMidlands0120to0820$Crime.type),
crimesoutcomesWM2020table[4]/length(crimesWestMidlands0120to0820$Crime.type),
crimesoutcomesWM2020table[5]/length(crimesWestMidlands0120to0820$Crime.type),
crimesoutcomesWM2020table[6]/length(crimesWestMidlands0120to0820$Crime.type),
crimesoutcomesWM2020table[7]/length(crimesWestMidlands0120to0820$Crime.type),
crimesoutcomesWM2020table[8]/length(crimesWestMidlands0120to0820$Crime.type),
crimesoutcomesWM2020table[9]/length(crimesWestMidlands0120to0820$Crime.type),
crimesoutcomesWM2020table[10]/length(crimesWestMidlands0120to0820$Crime.type),
crimesoutcomesWM2020table[11]/length(crimesWestMidlands0120to0820$Crime.type))
```





