---
title: "Oxford Government Response Tracker"
author: "Vanessa Fillis"
date: "03/11/2020"
output: html_document
---

to do
- rename variables/ dataframes
- better descriptions 


## Import the data

"The Oxford COVID-19 Government Response Tracker (OxCGRT) systematically collects information on several different common policy responses that governments have taken to respond to the pandemic on 18 indicators such as school closures and travel restrictions. It now has data from more than 180 countries." 

The data can be downloaded from GitHub: https://github.com/OxCGRT/covid-policy-tracker/tree/master/data. It is updated daily. 

(note: also data on timeseries)

```{r}
allcountries <- read.csv(url("https://github.com/OxCGRT/covid-policy-tracker/raw/master/data/OxCGRT_latest.csv"), stringsAsFactors = FALSE)
```

##Look at the latest data

Let's concentrate on the latest data available which is the day before today.

(Depending on the time of the day the map/ script is updated we could also directly look at today's data, so far data has been updated around 1pm)

```{r}
today <- lubridate::today()
yesterday <- today-1
```

##Format date column

There is one problem: the date in the original data frame is stored differently than the data in our new variable *yesterday*. In order to filter the *allcountries* data frame according to yesterday's date, we have to first format its date and save the changes in a new column.

```{r}
allcountries$Date.edited <- lubridate::ymd(allcountries$Date)
```

##Filter

Filter the data, so that only the latest entries (from the day before) remain. 

```{r}
allcountries_latestdata <- subset(allcountries, allcountries$Date.edited == yesterday)
``` 

##Remove unnecessary regions

For the visualisation in a world map we don't need data for regions within a country, e.g. for England, Northern Ireland, Scotland and Wales individually as the map displays the score for the entire United Kingdom. 

Therefore we will remove the rows that solely include data for regions. 

```{r}
allcountries_latestdata <- allcountries_latestdata[!(allcountries_latestdata$RegionCode=="UK_ENG" | allcountries_latestdata$RegionCode=="UK_NIR" | allcountries_latestdata$RegionCode=="UK_SCO" | allcountries_latestdata$RegionCode=="UK_WAL" | allcountries_latestdata$RegionCode=="US_WY" | allcountries_latestdata$RegionCode=="US_WV" | allcountries_latestdata$RegionCode=="US_WI" | allcountries_latestdata$RegionCode=="US_WA" | allcountries_latestdata$RegionCode=="US_VT" | allcountries_latestdata$RegionCode=="US_VI" | allcountries_latestdata$RegionCode=="US_VA" | allcountries_latestdata$RegionCode=="US_UT" | allcountries_latestdata$RegionCode=="US_TX" | allcountries_latestdata$RegionCode=="US_TN" | allcountries_latestdata$RegionCode=="US_SD" | allcountries_latestdata$RegionCode=="US_SC" | allcountries_latestdata$RegionCode=="US_RI" | allcountries_latestdata$RegionCode=="US_PA" | allcountries_latestdata$RegionCode=="US_OR" | allcountries_latestdata$RegionCode=="US_OK" | allcountries_latestdata$RegionCode=="US_OH" | allcountries_latestdata$RegionCode=="US_NY" | allcountries_latestdata$RegionCode=="US_NV" | allcountries_latestdata$RegionCode=="US_NM" | allcountries_latestdata$RegionCode=="US_NJ" | allcountries_latestdata$RegionCode=="US_NH" | allcountries_latestdata$RegionCode=="US_NE" | allcountries_latestdata$RegionCode=="US_ND" | allcountries_latestdata$RegionCode=="US_NC" | allcountries_latestdata$RegionCode=="US_MT" | allcountries_latestdata$RegionCode=="US_MS" | allcountries_latestdata$RegionCode=="US_MO" | allcountries_latestdata$RegionCode=="US_MN" | allcountries_latestdata$RegionCode=="US_MI" | allcountries_latestdata$RegionCode=="US_ME" | allcountries_latestdata$RegionCode=="US_MD" | allcountries_latestdata$RegionCode=="US_MA" | allcountries_latestdata$RegionCode=="US_LA" | allcountries_latestdata$RegionCode=="US_KY" | allcountries_latestdata$RegionCode=="US_KS" | allcountries_latestdata$RegionCode=="US_IN" | allcountries_latestdata$RegionCode=="US_IL" | allcountries_latestdata$RegionCode=="US_ID" | allcountries_latestdata$RegionCode=="US_IA" | allcountries_latestdata$RegionCode=="US_HI" | allcountries_latestdata$RegionCode=="US_GA" | allcountries_latestdata$RegionCode=="US_FL" | allcountries_latestdata$RegionCode=="US_DE" | allcountries_latestdata$RegionCode=="US_DC" | allcountries_latestdata$RegionCode=="US_CT" | allcountries_latestdata$RegionCode=="US_CO" | allcountries_latestdata$RegionCode=="US_CA" | allcountries_latestdata$RegionCode=="US_AZ" | allcountries_latestdata$RegionCode=="US_AR" | allcountries_latestdata$RegionCode=="US_AL" | allcountries_latestdata$RegionCode=="US_AK"),]
```

##Export

Export the data, so that it can be visualized, e.g. with Datawrapper or Flourish. 

```{r}
write.csv(allcountries_latestdata, file = "govresponse_allcountries_latest_data.csv")
```

##Create subsets

We can also create subsets for specific topics/ questions, for example for the different policy indices. 

These are the four indices: 

  - overall government response index (all indicators)
  - containment and health index (all C and H indicators)
  - stringency index (all C indicators, plus H1 which records public information         campaigns)
  - economic support index (all E indicators)

The indices aggregate the data into a single number. Each of these indices report a number between 0 to 100 that reflects the level of the governments response along certain dimensions. This is a measure of how many of the relevant indicators a government has acted upon, and to what degree. The index cannot say whether a government's policy has been implemented effectively.

##Government Response Index

```{r}
allcountries_latest_govresponse <- allcountries_latestdata[c(1:5,42,47)]
write.csv(allcountries_latest_govresponse, file = "all_latest_govresponse.csv")
```

##Containment and Health Index

```{r}
allcountries_latest_healthindex <- allcountries_latestdata[c(1:5,44,47)]
write.csv(allcountries_latest_healthindex, file = "all_latest_healthindex.csv")
```

##Stringency Index

```{r}
allcountries_latest_stringency <- allcountries_latestdata[c(1:5,38,47)]
write.csv(allcountries_latest_stringency, file = "all_latest_stringency.csv")
```

##Economic Support Index

```{r}
allcountries_latest_economicsupport <- allcountries_latestdata[c(1:5,46,47)]
write.csv(allcountries_latest_economicsupport, file = "all_latest_economicsupport.csv")
```


